---
title: "p8105_hw6_xy2517"
author: "Xuehan Yang"
date: "2021/12/2"
output: github_document
---

```{r, message=FALSE}
library(tidyverse)
library(modelr)
library(viridis)
library(GGally)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Problem 1

Explore the effects of several variables on a child’s birthweight.

### Load and clean data for regression

```{r, factor variable, message=FALSE, warning=FALSE}
bwt_df = read_csv("./data/birthweight.csv") %>% 
  mutate(
    babysex = as_factor(babysex),
    babysex = fct_recode(babysex, "male" = "1", "female" = "2"),
    frace = as_factor(frace),
    frace = fct_recode(frace, "White" = "1", "Black" = "2", "Asian" = "3",
                        "Puerto Rican" = "4", "Other" = "8", "Unknown" = "9"),
    malform = as_factor(malform),
    malform = fct_recode(malform, "absent" = "0", "present" = "1"),
    mrace = as_factor(mrace),
    mrace = fct_recode(mrace, "Whie" = "1", "Black" = "2", "Asian" = "3",
                        "Puerto Rican" = "4", "Other" = "8"))
```

```{r, check NA}
which(is.na(bwt_df))
```
There is no NA value.

```{r, check missing values}
bwt_df %>% 
  filter(bhead == 0 | blength == 0 | bwt == 0 | delwt == 0 | fincome == 0 | gaweeks == 0 | menarche == 0 | mheight == 0 |
           momage == 0 | ppbmi == 0 | wtgain == 0) %>% 
  knitr::kable()
```
As for wt_gain, it is likely that mother gains no weight during pregnancy.

As for family income, without any income is also possible. 

Menarche should not be zero because it is the mother's age at menarche.

Parity, pnumlbw, pnumsga these three variables contain two many zeros, which is not good for regression, so they should not be included in the model.

```{r}
ggplot(bwt_df, aes(x = smoken)) + geom_histogram(bins = 10) +
  labs(
    x = "average number of cigarettes smoked per day during pregnancy",
    y = "Count",
    title = "Distribution of average numberof cigarettes smoking"
  )
```

Because there are so many nonsmokers during pregnancy, I want to transfer the smoken variable into category.

smoken: smoking status(0 = Non(0), 1 = Light(0-10), 2 = Heavy(10+))

```{r}
bwt_df = bwt_df %>% 
  filter(bhead != 0 & blength != 0 & bwt != 0 & delwt != 0 & gaweeks != 0 & menarche != 0 & mheight != 0 & momage != 0 & ppbmi != 0) %>% 
  select(-parity, -pnumlbw, -pnumsga) %>% 
  mutate(
  smoken = case_when(
    smoken == 0 ~ "Non",
    smoken > 0 & smoken <=10 ~ "Light",
    smoken >10 ~ "Heavy"
  )
) %>% 
  mutate(
    smoken = as_factor(smoken),
    smoken = fct_relevel(smoken, c("Non", "Light", "Heavy"))
    ) %>% 
  relocate(bwt, babysex, frace, mrace, malform, smoken) # put categorical variables forward
```

Final data set after cleaning
```{r}
head(bwt_df) %>% 
  knitr::kable()
```

There are total `r nrow(bwt_df)` observations of babies' weight, with `r ncol(bwt_df)` variables. Variables contain `r names(bwt_df)`

### Propose a regression model

**Potential predictors**

* babysex: Birth weight has consistently been shown to be higher in boys than in girls.(ref1)
* smoken: Smoking has been shown to affect glucose homeostasis and increase the risk for type 2 diabetes mellitus. Further, gestational diabetes mellitus (GDM) and smoking are known to influence offspring birthweight.(ref2)
* fincome: A clear graded association between income quintile and low birth weight was apparent in the United States(ref3)
* gaweeks: There was a varied difference of distribution in birth weights for gestational age.(ref4)
* wtgain: Excessive pregnancy weight gain appears to increase birth weight(ref5)
* ppbmi: Women with abnormal pre-pregnancy body mass index are at risk for adverse birth weight outcomes.(ref6)
* ppwt: Women with abnormal pre-pregnancy body mass index are at risk for adverse birth weight outcomes.(ref6)
* bhead: head is a part of birthweight.
* blength: genrally, longer length bigger weight.
* mheight: A bigger mom tends to deliver a bigger baby.
* momage: different ages indicate different health status, which might influence the birthweight of baby.

**Look at the relation between birthweight and predictors**

```{r}
ggplot(bwt_df, aes(x = babysex, y = bwt)) + 
  geom_boxplot() +
  labs(
    x = "Sex",
    y = "Birthweight",
    title = "Association between Birthweight and Sex"
  )
ggplot(bwt_df, aes(x = smoken, y = bwt)) + 
  geom_boxplot() +
  labs(
    x = "Smoking Status",
    y = "Birthweight",
    title = "Association between Birthweight and smoking status"
  )
```

Male babies tend to have higher birthweight. Smoking tend to lower the birthweight of baby.

```{r,fig.asp = 1, fig.width=8}
corr_df = bwt_df %>% 
  select(-babysex,-frace,-mrace,-malform,-smoken)

ggpairs(corr_df, 
        lower = list(continuous = wrap("points", size = 0.01)), 
        upper = list(continuous = wrap("cor", size = 2.5))) + 
  theme(axis.text.x = element_text(size = 6, angle = 45),
        axis.text.y = element_text(size = 6, angle = 45))
```

We can see there is colinearity between ppbmi and ppwt.(correlation coefficient > 0.7). So we keep one of them, ppbmi.

Final prediction dataset.
```{r}
predict_df = bwt_df %>% 
  select(-frace, -mrace, -malform, -ppwt)
```

**Regression model**

```{r}
bwt_mod = lm(data = bwt_df, bwt ~ babysex + smoken + bhead + blength + delwt + fincome + gaweeks + menarche + mheight + momage + ppbmi + wtgain)
bwt_mod %>% broom::tidy() %>% knitr::kable(digits = 3)
```

Delete the insignificant variables, delwt, mheight, ppbmi, wtgain

```{r}
bwt_mod = lm(data = bwt_df, bwt ~ babysex + smoken + bhead + blength + fincome + gaweeks + menarche + momage)
bwt_mod %>% 
  broom::tidy() %>%
  select(term, estimate, p.value) %>% 
  mutate(
    term = str_replace(term, "smoken", "smo_status: "),
    term = str_replace(term, "babysex", "sex: ")) %>% 
  knitr::kable(digits = 3)
```

All the predictors are significant. Adjusted R square is `r bwt_mod %>% broom::glance() %>% select(adj.r.squared)`, which means 69.4% variation in birthweight can be explained by predictors. Female babies are on average 32.87 grams heavier than male babies. Each centimeter increase in baby's head is associated with 138.1 grams increase in birthweight. Birthweights of baby delivered by lightsmokers and highsmokers during pregnancy are on average 23.8 grams and 46.9 grams less than nonsmokers. 

**Diagnostic**

```{r}
bwt_df %>% 
  modelr::add_predictions(bwt_mod) %>% 
  modelr::add_residuals(bwt_mod) %>% 
  ggplot(aes(x = pred, y = resid)) + geom_point()
```

There are some potential outliers on the left. And it seems there exists heteroscedasticity as the variance of resid increase with hte rising of predicted birthweight.

**Compare with other two models**

```{r}
cv_df = 
  crossv_mc(bwt_df,100)

cv_df = 
  cv_df %>% 
  mutate(
    train = map(train, as_tibble),
    test = map(train, as_tibble)
  ) %>% 
  mutate(
    bwt_mod = map(train, ~lm(bwt ~ babysex + smoken + bhead + blength + fincome + gaweeks + menarche + momage, data = .x)),
    model_2 = map(train, ~lm(bwt ~ blength + gaweeks, data = .x)),
    model_3 = map(train, ~lm(bwt ~ bhead + blength + babysex + bhead*blength + bhead*babysex + blength*babysex, data = .x))
  ) %>% 
  mutate(
    rmse_mymodel = purrr::map2(bwt_mod, test, ~rmse(model = .x, data = .y)),
    rmse_maineffect = purrr::map2(model_2, test, ~rmse(model = .x, data = .y)),
    rmse_interaction = purrr::map2(model_3, test, ~rmse(model = .x, data = .y))
  )
```

```{r}
cv_df %>% 
  select(starts_with("rmse")) %>% 
  unnest(cols = c(rmse_mymodel, rmse_maineffect, rmse_interaction)) %>% 
  pivot_longer(
    everything(),
    names_to = "model",
    values_to = "rmse",
    names_prefix = "rmse_"
  ) %>% 
  mutate(model = fct_reorder(model, rmse)) %>% 
  ggplot(aes(x = model, y = rmse)) + geom_boxplot() +
  labs(
    x = "Model",
    y = "RMSE",
    title = "Comparison of three models"
  )
```

Model 1 has the minimum rmse, so I pick model 1 as the final model for prediction. 

## Problem 2

Data here.
```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```

Funciton to draw one bootstrap sample.

```{r}
boot_sample = function(df){
  sample_frac(df, replace = TRUE)
}
```

Bootstrapping 5000 samples.

```{r}
boot_straps = 
  tibble(
    number = 1:5000,
    sample = rerun(5000, boot_sample(weather_df))
  )
```

SLR and extracting the two quantities into estimate_df
```{r, warning=FALSE   }
boot_straps =
  boot_straps %>% 
  mutate(
    model = map(sample, ~lm(tmax~tmin, data = .x)),
    tidy = map(model, broom::tidy),
    glance = map(model, broom::glance))

estimate_df = 
  boot_straps %>%
  unnest(tidy) %>%
  select(number,glance, term, estimate) %>%
  mutate(
    term = str_replace(term, "(Intercept)", "beta_zero"),
    term = str_replace(term, "tmin", "beta_one")
  ) %>% 
  pivot_wider(
    names_from = term,
    values_from = estimate
  ) %>% 
  unnest(glance) %>%
  janitor::clean_names() %>% 
  mutate(quantity_log = log(beta_one*beta_zero)) %>% 
  select(number, quantity_log, r_squared)
```

Plot the distribution of estimates

```{r}
estimate_df %>% 
  ggplot(aes(x = r_squared)) + geom_density() +
  labs(
    x = "r_squared",
    y = "density"
  )
```

The distribution of r_squared is nearly normal around 0.91.

```{r}
quantile(estimate_df %>% pull(r_squared), prob = c(0.025, 0.975))
```

The 95% confidence interval for r_squared is (0.893, 0.928)

```{r}
estimate_df %>% 
  ggplot(aes(x = quantity_log)) + geom_density() +
  labs(
    x = "log(beta0*beta1)",
    y = "density"
  )
```

The distribution of log(beta0*beta1) is nearly normal around 2.02.

```{r}
quantile(estimate_df %>% pull(quantity_log), prob = c(0.025, 0.975))
```

The 95% confidence interval for log(beta0*beta1) is (1.964, 2.059)

## Reference

1.Van Vliet, Guy; Liu, Shiliang; Kramer, Michael S. Decreasing Sex Difference in Birth Weight, Epidemiology: July 2009 - Volume 20 - Issue 4 - p 622
doi: 10.1097/EDE.0b013e3181a82806 

2.Masalin, S, Kautiainen, H, Gissler, M, Pennanen, P, Eriksson, JG, Laine, MK. Impact of smoking on gestational diabetes mellitus and offspring birthweight in primiparous women. Acta Obstet Gynecol Scand. 2020; 99: 1632– 1639. https://doi.org/10.1111/aogs.13924

3.Melissa L. Martinson and Nancy E. Reichman, 2016:
Socioeconomic Inequalities in Low Birth Weight in the United States, the United Kingdom, Canada, and Australia
American Journal of Public Health 106, 748_754, https://doi.org/10.2105/AJPH.2015.303007

4.Jean Dupont Kemfang Ngowa, Irénée Domkam, Anny Ngassam, Georges Nguefack-Tsague, Walter Dobgima Pisoh, Cyrille Noa, Jean Marie Kasia, "References of Birth Weights for Gestational Age and Sex from a Large Cohort of Singleton Births in Cameroon", Obstetrics and Gynecology International, vol. 2014, Article ID 361451, 8 pages, 2014. https://doi.org/10.1155/2014/361451

5.Ludwig DS, Currie J. The association between pregnancy weight gain and birthweight: a within-family comparison. Lancet. 2010;376(9745):984-990. doi:10.1016/S0140-6736(10)60751-9

6.Zhao R, Xu L, Wu ML, Huang SH, Cao XJ. Maternal pre-pregnancy body mass index, gestational weight gain influence birth weight. Women Birth. 2018 Feb;31(1):e20-e25. doi: 10.1016/j.wombi.2017.06.003. Epub 2017 Jul 14. PMID: 28716548.

